---
title: "DV-04"
author: "Kiran Nagar"
format: html
embed-resources: true
editor: visual
---

# DV-04

## Question 1

We read in the data as instructed.

```{r}
#| message: false
library(tidyverse)

# Read in the data 
tv_ratings <- read_csv("https://raw.githubusercontent.com/vaiseys/dav-course/main/Data/tv_ratings.csv")

# Glimpse the data 
glimpse(tv_ratings)
```

```{r}
tv_long <- tv_ratings %>% 
  group_by(title) %>% 
  summarise(num_seasons = n()) %>% 
  ungroup() %>% 
  left_join(tv_ratings, by = "title") 

tv_long <- tv_long %>% 
  filter(num_seasons >= 5)
```

```{r}
p1 <- ggplot(tv_long,
             aes(x = seasonNumber,
                 y = av_rating,
                 group = title)) + geom_line()
p1
```

It is too messy to draw any meaningful conclusions due to the messiness other than the fact that not very many shows go 15+ seasons.

## Question 2

```{r}
p1faceted <- ggplot(tv_long,
             aes(x = seasonNumber,
                 y = av_rating,
                 group = title)) + geom_line() + facet_wrap(~genres)
p1faceted
```

The shows that last longer tend to include any of the Crime, Mystery, and Drama genre classifications. In general, there doesn't tend to be a huge amount of ratings fluctuation for shows---that is, good shows tend to stay good-ish, and bad shows don't often turn things around. There are a couple of examples where this is not true, such as *Are You Afraid of the Dark?*, which was labeled with `Drama, Family, and Fantasy`.

## Question 3

```{r}
top_rated <- tv_ratings %>%
  filter(av_rating >= 9)

genreplot <- ggplot(top_rated, aes(x = genres,)) + geom_bar()
genreplot
```

```{r}
genreplot + coord_flip()
```

`coord_flip()` makes the x and y-axes swap places, which helps for readability here. The genre with the most top-rated shows is `Drama`.

## Question 4

```{r}
comedies_dramas <- tv_ratings %>% 
  mutate(is_comedy = if_else(str_detect(genres, "Comedy"), 
                             1, 
                             0)) %>% # If it contains the word comedy then 1, else 0
  filter(is_comedy == 1 | genres == "Drama") %>% # Keep comedies and dramas
  mutate(genres = if_else(genres == "Drama", # Make it so that we only have those two genres
                          "Drama", 
                          "Comedy"))

glimpse(comedies_dramas)
```

```{r}
library(patchwork)
```

```{r}
comedies <- comedies_dramas %>%
  filter(is_comedy == 1)
dramas <- comedies_dramas %>% 
  filter(is_comedy == 0)

p1 <- ggplot(comedies,
             aes(x = av_rating)) +
  geom_density() +
  labs(title = "Comedies",
       x = "Average Rating",
       y = "Density")

p2 <- ggplot(dramas,
             aes(x = av_rating)) +
  geom_density() +
  labs(title = "Dramas",
       x = "Average Rating",
       y = "Density")

p1 + p2



p3 <- ggplot(comedies_dramas,
             aes(x = av_rating,
                 group = genres,
                 color = genres)) +
  geom_density() +
  labs(title = "Comedies v. Dramas",
       x = "Average Rating",
       y = "Density")

p3
```

I think the prediction holds to be somewhat-true. There is a much higher density of comedies around the 8 rating, but above that, dramas win out in terms of density between scores of 8-10.

## Question 5

```{r}
p3 <- ggplot(comedies_dramas,
             aes(x = av_rating,
                 group = genres,
                 color = genres)) +
  geom_histogram() +
  labs(title = "Comedies vs. Dramas",
       x = "Average Rating",
       y = "Count")
p3
```

```{r}
p4 <- ggplot(comedies_dramas,
             aes(x = av_rating,
                 group = genres,
                 color = genres)) +
  geom_freqpoly()
p4
```

I notice that there is a very close relationship between the `geom_freqpoly()` and the `geom_histogram()`, which makes sense since they are both based on the count within certain bins. Of the three plots, I think that the `geom_freqpoly()` is most useful because it's intuitive and it gives you an idea for the number of shows that correspond to a given rating.

## Question 6

```{r}
share_plot <- ggplot(comedies_dramas,
                     aes(x = av_rating,
                         y = share,)) +
  geom_bin2d()

share_plot
```

This plot indicates to me that there are a ton of shows which received a pretty good rating (\~8) and still got almost no share of the audience. It is also notable that the highest rated shows (between 9-10) don't receive an increase in share. This plot clearly illustrates the density of points in an area on the graph which a scatter plot struggles to do.

```{r}
genre_share_plot <- ggplot(comedies_dramas,
                     aes(x = av_rating,
                         y = share,
                         fill = genres)) +
  geom_bin2d()
genre_share_plot
```

From this plot, I noticed that the comedy series tend to get a larger audience share. This makes sense to me, since I think people like to watch series that make them laugh! It looks like the outlier is Season 1 of *Dekalog*, a drama released in 1990. Curiously, it was not renewed for a second season!
