---
title: "HW 3"
author: "Kiran Nagar"
format: html
embed-resources: true
editor: visual
---

```{r}
library(tidyverse)

mario_kart <- read_csv("https://raw.githubusercontent.com/NicolasRestrep/223_course/main/Data/world_records.csv")
glimpse(mario_kart)
```

```{r}
three_laps <- mario_kart %>% filter(type == "Three Lap")
no_rainbow <- three_laps %>% filter(track != "Rainbow Road")
rainbow_records <- three_laps %>% filter(track == "Rainbow Road")
```

## Question 2

```{r}
rainbow_stats <- rainbow_records %>% 
  summarize(
    mean = mean(time, na.rm = TRUE),
    std_dev = sd(time, na.rm = TRUE)
  )

other_tracks_stats <- no_rainbow %>% 
  summarize(
    mean = mean(time, na.rm = TRUE),
    std_dev = sd(time, na.rm = TRUE)
  )

```

The mean time at Rainbow Road was more than double that that of the other tracks---the times were 275.63s and 113.80s, respectively. The standard deviation of Rainbow Road was also almost double that of the other tracks, at 91.82s versus 52.98s.

## Question 3

```{r}
different_records <- three_laps %>% 
  group_by(track) %>% 
  summarize(count = n()) %>% 
  arrange(-count)

top_track <- different_records[1,]
```

Toad's Turnpike has the most records established in it.

## Question 4

```{r}
player_track_records <- three_laps %>% 
  group_by(player, track) %>% 
  summarize(count = n()) %>% 
  arrange(-count)

top_driver <- player_track_records[1,]
```

Penev has recorded the most wins at any one track. That track is Choco Mountain, where he has won a whopping 26 times.

## Question 5

```{r}
track_averages <- three_laps %>% 
  group_by(track) %>% 
  summarize(avg_time = mean(time, na.rm = TRUE)) %>%
  arrange(-avg_time)
```

Rainbow Road has the highest average time at 275.63s.

```{r}
track_bests <- three_laps %>% 
  group_by(track) %>% 
  arrange(time) %>% 
  slice(1) %>%
  select(track, time)

```

## Question 6

```{r}
three_laps <- three_laps %>% 
  mutate(long_duration = if_else(record_duration > 100, 1, 0))

```

```{r}
player_long_duration <- three_laps %>% 
  group_by(player) %>% 
  summarize(total_long_duration = sum(long_duration)) %>% 
  arrange(-total_long_duration)
```

MR has the most long duration records, with 81.

## Question 7

```{r}
drivers <- read_csv("https://raw.githubusercontent.com/NicolasRestrep/223_course/main/Data/drivers.csv")
```

```{r}
three_laps_joined <- left_join(three_laps, drivers, by = "player")
country_records_plot <- three_laps_joined %>% 
  group_by(nation) %>% 
  tally() %>% 
  ggplot(aes(x = reorder(nation, n), y = n)) +
  geom_bar(stat = "identity") +
  coord_flip()
country_records_plot
```
