---
title: "HW 4"
author: "Kiran Nagar"
format: html
embed-resources: true
editor: visual
---

```{r}
#| message: false
library(tidyverse)
```

## Question 1

```{r}
#| message: false
nfl_data <- read_csv("https://raw.githubusercontent.com/NicolasRestrep/223_course/main/Data/nfl_salaries.csv")

```

Dataset was read in from the link as instructed.

## Question 2

```{r}
tidy_nfl_data_pivot <- nfl_data %>%
  pivot_longer(cols = "Cornerback":"Wide Receiver",
               names_to = "position",
               values_to = "salary in $M")
tidy_nfl_data_pivot <- tidy_nfl_data_pivot %>% 
  mutate(`salary in $M` = (`salary in $M`/(10^6)))
tidy_nfl_data_pivot

```

Dataset was reshaped so that every row tells us the salary associated with a position in a certain year.

## Question 3

```{r}
qbs_only <- tidy_nfl_data_pivot %>% filter(position == 'Quarterback')

ggplot(qbs_only, aes(x = `salary in $M`)) + geom_histogram(bins = 10)
```

```{r}
ggplot(qbs_only, aes(x = `salary in $M`)) + geom_histogram(bins = 15) + facet_wrap(vars(year)) + labs(y = "Number of Players")
```

I notice that there are a lot of quarterbacks who earn on the low end of the distribution every year, and then a few outliers earning at a high level. This makes sense to me, because every team has two backup quarterbacks earning a small salary, and one main option earning at a high level.

## Question 4

```{r}
average_salary_millions <- tidy_nfl_data_pivot %>% 
  group_by(position, year) %>% 
  summarize(average_in_millions = mean(`salary in $M`, na.rm = TRUE))
average_salary_millions
```

I used `group_by` and `summarize` to create a new dataset called `average_salary`.

## Question 5

```{r}
ggplot(average_salary_millions, aes(x = year, y = average_in_millions)) + geom_line() + facet_wrap(vars(position))+ theme_minimal() + labs(y = "Average Salary in $M", x="Year")
```

I generated the linegraph as requested in the assignment. One trend that is apparent to me is the somewhat stair-stepping nature of the QB position. This makes sense to me, as historically every couple of years the league-leading QB signs a contract which pushes the whole market higher. Running backs have also earned at a level that has stayed pretty similar, even declining a bit in recent years. This makes sense to me, since we've recently seen RBs holding out for better salaries, and GMs not being willing to pay RBs with "tread on their tires."
