---
title: "HW 1"
author: "Kiran Nagar"
format: html
embed-resources: true
editor: visual
---

## Question 1

I installed the three packages in the console, as I was instructed to do.

## Question 2

```{r}
library(causact)
library(dplyr)
library(igraph)

df <- dplyr::as_data_frame(x = c(1,2,3))
glimpse(df)

data <- dplyr::n_distinct(x <- c(5,6,2,7,9,1))
```

My hypothesis is that since the function names are global, whichever package is installed most recently would overwrite the global function's namespace. In this case, `igraph` is installed after `dplyr`, so R called `igraph`'s `as_data_frame` instead of `dplyr`'s.

## Question 3

```{r}
?n_distinct()
```

In a set of numbers (or vector), `n_distinct()` returns the number of unique combinations that can be made. It does not count a pair in reverse order as two pairs, as implied by the name. `n_distinct()` can handle null set vectors, though it also has support to filter them out.

## Question 4

```{r}
glimpse(baseballData)
```

The dataset has 12,145 rows and 5 columns. The `Home` column is of type `<fct>`, which means it is a factor variable. This basically means it defines a category. The `HomeScore` variable is of type `<int>`, which is a quantitative variable that only takes integer values.

## Question 5

```{r}
baseballData[1,]
```

One row in the data represents one baseball game.

```{r}
baseballData[,2:3] %>% head()
```

These two columns represent the two teams playing in MLB games from history, as well as who was the home team and who was the visitor. Without including the first column (Date), this information is not particularly useful.

## Question 6

```{r}
name <-
  c(
    "Wayne Gretzky",
    "Gordie Howe",
    "Jaromir Jagr",
    "Brett Hull",
    "Marcel Dionne",
    "Phil Esposito" ,
    "Mike Gartner",
    "Alex Ovechkin",
    "Mark Messier" ,
    "Steve Yzerman"
  )

goals <- c(894, 801, 766, 741, 731, 717, 708, 700, 694, 692)

year_started <- c(1979, 1946, 1990, 1986, 1971, 1963, 1979, 2005, 1979, 1983)

```

```{r}
df <- tibble( 
  Name = name, 
  Goals = goals,
  Year_Joined = year_started ) 

glimpse(df)
```
