---
title: "HW-MD0809"
author: "Kiran Nagar"
format: html
embed-resources: true
editor: visual
editor_options: 
  chunk_output_type: console
---

# HW-MD0809

```{r}
#| message: false
library(tidyverse)
library(here)
library(moderndive)
library(infer)
theme_set(theme_minimal())
```

# Chapter 8

## Question 1

```{r}
set.seed(1108)

poll <- tibble(
  vote_gop = rbinom(n = 1000,
                    size = 1,
                    prob = .53))
```

```{r}
# Bootstrap resampling
set.seed(1108)  # Setting seed for reproducibility
bootstrap_dist <- poll %>%
  rep_sample_n(size = nrow(poll), replace = TRUE, reps = 1000) %>%
  group_by(replicate) %>%
  summarize(mean_vote = mean(vote_gop))

# Calculating 95% Confidence Interval
ci_one_way <- quantile(bootstrap_dist$mean_vote, c(.025, .975))
print(ci_one_way)
```

Using this first method, I got a percentile-based confidence interval of (0.483975, 0.548000).

```{r}
# Infer approach
set.seed(1108)  # Setting seed for reproducibility
ci_second_way <- poll %>%
  specify(response = vote_gop) %>%
  generate(reps = 1000, type = "bootstrap") %>%
  calculate(stat = "mean") %>%
  get_ci(level = 0.95, type = "percentile")

print(ci_second_way)
```

Using the second method, I got a percentile-based confidence interval of (0.483975, 0.548).

The two confidence intervals are identical. If you don't set the same seed, then the confidence intervals will not be the same. This is because when "randomly" sampling, setting a different seed would cause the computer to generate randomness differently.

## Rap Data

```{r}
rap_poll <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-04-14/polls.csv")
```

```{r}
top_songs <- rap_poll %>% 
  filter(rank == 1)
```

## Question 2

```{r}
years <- ggplot(top_songs,
                aes(x = year)) +
  geom_histogram(binwidth = 1,
                 center = 0) +
  labs(title = "Release Year of Critics' #1 Favorite Rap Songs",
       x = "Release Year",
       y = "Number of Songs")
years
```

1994 was the year with the most commonly named favorite track in the critic poll, with 14 critics naming their #1 from that year. It doesn't hurt that "Juicy" by Biggie was released in 1994!

## Question 3

```{r}
set.seed(3)
boot_d <- top_songs %>% 
  specify(response = year) %>% 
  generate(reps = 1000, type = "bootstrap") %>% 
  calculate(stat = "mean")
```

```{r}
confidence_interval <- boot_d %>% 
  get_ci(level = 0.95, type = "percentile")
print(confidence_interval)
```

The lower bound is 1993, and the upper bound is 1996.

## Question 4

```{r}
set.seed(3)
boot_d2 <- rap_poll %>% 
  rep_sample_n(size = 25, reps = 1000, replace = TRUE) %>% 
  group_by(replicate) %>% 
  summarize(mean_year = mean(year))

ci_25 <- quantile(boot_d2$mean_year, c(.025, .975))
print(ci_25)
```

This generated a very different confidence interval for me, with a lower bound of 1995 and an upper bound of 2002.

# Chapter 9

```{r}
pl_data <- read_csv("https://raw.githubusercontent.com/vaiseys/dav-course/main/Data/premier_league.csv")

glimpse(pl_data)
```

## Question 5

```{r}
home_wins <- pl_data %>% 
  filter(result == "hw")

home_wins_prop <- nrow(home_wins) / nrow(pl_data)

print(home_wins_prop)
```

The proportion of home wins during the 2015/2016 season was 0.4131579, or 41.3%.

## Question 6

```{r}
set.seed(22)

sampled_proportions <- c()

for (i in 1:1000) {
  
  sampled_results <- sample(c("aw", "hw" , "d"), 
                            size = 380,
                            replace = TRUE, 
                            prob = c(1/3,1/3,1/3))
  prop <- sum(sampled_results == "hw")/380
  sampled_proportions[i] <- prop
  
}
proportions_df <- data.frame(Proportion = sampled_proportions)
```

```{r}
proportions_graph <- ggplot(proportions_df,
                            aes(x = Proportion)) + 
  geom_histogram(binwidth = 0.01, boundary = 0.1, color = "white") +
  labs(title = "Sample Proportions with Equal Probability Results",
       x = "Proportion",
       y = "Number of Proportions")
proportions_graph
```

The histogram pretty closely resembles a normal distribution, which makes sense since we are taking a huge number of samples, which by the CLT means that the distribution will get closer and closer to a normal distribution. The proportion in Question 5 was much higher than the sample proportions in Question 6. This makes total sense, because in real life, a home win is more likely than an away win.

## Question 7

In this scenario, the null hypothesis would be that Home/Away has no bearing, and the Home team and the Away team win at the same rate. The alternate hypothesis would be that the Home and Away teams win at different rates.

## Question 8

In our scenario, with a two-sided hypothesis test, the p-value is the probability that the proportion of home wins is just as or more extreme than the proportion of home wins when we assume all three possible results are equally probable. Visually, it would be two shaded portions on the right and left sides of the null distribution of home win proportions.
